%----------------------------------------------------------------------% 
% 2002 - Q4

(a)  Nine per cent of PCB boards purchased over the Internet from the Far East have some defect.  From a large consignment of boards, 50 are chosen at random.  What is the probability that:
(i)5 or more boards have some defect?
(ii)Exactly 3 boards are defective?
(iii)Less than 2 boards are defective?
(iv)If more than 8 boards from the 50 were defective what action would you take? (justify)

%------------------------------------------%

(b) Flaws occur in a heavy duty industry carpet at the rate of 15 per 10 square metre section.  Calculate the probability that:
(i)3 or more flaws will occur in a one square metre section
(ii)Exactly 4 flaws will occur in a 10 metre square section
(iii)14 or less flaws will occur in a 20 square metre section

(c)There is a constant probability of 0.10 that the power supply in telecoms network will not start.  You are requested to calculate the probability that the power supply will fail the 5th time it is activated.%-----------------------------------------------------------------------------% 
% 2001 - Q 3

(a) A new printer is fully warranted during its first year. The number of service calls  during the year of warranty is Poisson distributed with a mean of 1.0. 

\begin{itemize} 
	\item[(i)] What is the probability of no service calls during the warranty period for a printer?                                                                                                            4 
	
	\item[(ii)] What is the probability of fewer than 3 service calls? 4 
	
	\item[(iii)] If each service call costs £20, what is the expected cost of warranty service calls per  printer? What is the standard deviation of the cost? What is the probability that the  total cost of service calls for the year will exceed the expected cost? 4 
	
	
	\item[(iv)] If it were discovered that 45\% of printers had no warranty service calls but about 9\%   required 4 or more calls, would this correspond with the results of a Poisson process?  Explain. 4 
\end{itemize}

%----------------------------------------------------------------------% 
(b) A manufacturer of radios claims that 5% of his radios do not meet the quality 
specification. A shop keeper will accept a shipment if, one lot containing 10 radios, taken at random, contains only 0 or 1 defective ones; otherwise he will return the shipment. 

\begin{itemize} 
	\item[(i)] What is the probability that he will return the shipment when the  manufacturer's claim  is correct. 4  
	
	\item[(ii)] What is the probability that he will accept the shipment when there are actually 10\%  defectives? 5 
\end{itemize}


