\documentclass[]{report}

\voffset=-1.5cm
\oddsidemargin=0.0cm
\textwidth = 480pt

\usepackage{framed}
\usepackage{subfiles}
\usepackage{graphics}
\usepackage{newlfont}
\usepackage{eurosym}
\usepackage{amsmath,amsthm,amsfonts}
\usepackage{amsmath}
\usepackage{color}
\usepackage{amssymb}
\usepackage{multicol}
\usepackage[dvipsnames]{xcolor}
\usepackage{graphicx}
\begin{document}
%----------------------------------------------------------------%


%-------------------------------- %
\section{Probability Questions}





%===============================================%




The table below gives the number of thunderstorms reported in a particular summer
month by 100 meteorological stations.

\begin{tabular}{|c|c|c|c|c|c|c|}
	Number of thunderstorms: &0& 1 &2 &3 &4 &5 \\ \hline
	Number of stations: & 22 & 37 & 20 & 13 & 6 & 2 \\ \hline
\end{tabular}

\begin{itemize}
	\item[(a)] Calculate the sample mean number of thunderstorms.
	\item[(b)] Calculate the sample median number of thunderstorms.
	\item[(c)] Comment briefly on the comparison of the mean and the median.
\end{itemize}
%------------------------------------------------%
% 12 Normal Approximation of Binomial
% PMS April 2006 
In a certain large population 45\% of people have blood group A. 
A random sample of 300 individuals is chosen from this population.
Calculate an approximate value for the probability that more than 115 of the sample
have blood group A.

If X is the number in the sample with group A, then X has a binomial (300, 0.45)
distribution, so
\[ E[X] = 300 \times 0.45 = 135 \] and 
\[ Var[X] = 300 \times 0.45 \times 0.55 = 74.25 \].
Then, using the continuity correction,

\[ P(X > 115) = P(X > 115.5)\]
\[ 1- \frac{115.5 - 135}{\sqrt{74.25} } \]

%=========================================================% 
% PMS Autumn 2008 Question 1
The mean of a sample of 30 claims is \$5,200.
Six have mean of \$ 8000 (i.e. group 1)
Ten have mean of \$ 3100 (i.e. group 2)

Compute the mean for the remaining claims

\[\mbox{Total Costs} = (Cost for Group 1) + (Cost for Group 2) + (Cost from Group 3)\]

\begin{itemize}
	\item Total Cost for all three groups : $\$5200 \times 30 = \$156000$
	\item Cost for Group 1 : $\$8000 \times 6 = \$48000$
	\item Cost for Group 2 : $\$3100 \times 10 = \$31000$
\end{itemize}

Necessarily the cost for group 3 is $\$77000$

The mean claim for group 3 is therefore

\[ \frac{\$77000}{14} = \$5500 \]

%=========================================================% 

% PMS SPring 2006 Question 6
Poisson/Binomial/Exponential

\begin{itemize}
	\item  Poisson
	
	Find P(X=0) for Poisson Mean (m=0.5)
	
	
	\[ P(X=0) = \frac{e^{-0.5}}{0!}  = 0.606 \]
	
	
	%=================%
	\item Binomial
	
	%	\[ { 3 \choose 1} \times (0.606)^2 \times (1-0.606)^1 	= 0.434 \]
	
	
	
	%=================%
	\item Exponential
	
	No Claim in the next two years
	\[	= (0.606)^2 = 0.368\]
	
	
	\item 	Time Until Next Claim
	
	$\mu= 0.5$
	
	$T \approx exp(0.5)$
	
	\item	$P(XT >2) = exp(-1) = 0.368$
	
\end{itemize}



%http://www.profsullivan.com/storage/chapter%204%20notes.pdf
% http://mi.eng.cam.ac.uk/~rwp/Maths/eg/2012_maths_p9.pdf




%======================================================== %

\subsection{Descriptive Statistics - Question 4}
The following data give the marks of 10 students in a test (out of 20 marks). Calculate
\begin{itemize}
	
	\item[i)] the median    
	\item[ii)] the mean     
	\item[iii)] the range    
	\item[iv)] the standard deviation 
	\item[v)] The Inter-Quartile Range
\end{itemize}
\[12, 17, 7, 11, 18, 6, 14, 15, 11, 9.\]

%--------------------------------------------------------------------%


\subsection{Conditional Probability - VaVa-Voom}
VaVaVoom
The plant in Austria produces 80\% of the cars.
The plant in Belgium produces 20\% of the cars.
A randomly chosen car was build at Austrian plant
A randomly chosen car was built at the Belgian plant
S: A randomly chosen car has standard
The mean and standard deviation of the following

We are told the following piece of information $\bar{x} = 44$
So what is the coefficient of determination?

%==============================%
The duration, in months, of the construction phase of a number of motorway projects
were collected and tabulated as follows
Calculate the mean value of the durations.
Calculate the variance of the data set.
Calculate the standard deviation.
Calculate the coefficient of variation.
Firstly, what is the sample size?

\[cv= {s \over \bar{x}} \times 100\%\]

\[ \bar{x} = {49 + 55 + 43 + 45 + 41 +33 + 42 \over 7} \]
\[\bar{x} = 44\]
\[s^2 = 47\]
\[s = \sqrt{47}\]

\[{(49-44)^2 + (55-44)^2 + (43-44)^2 + (45-44)^2 + (41-44)^2 + (33-44)^2 + (42-44)^2 \over 7-1}\]
\[
25 + 121 + 1 + 1 + 9 + 121 + 4
={282/6} = 47
s = \sqrt{47}\]

= 15.58\%
%----------------------------------------------%
\subsection{Binomial Distribution}
According to a recent poll, approximately seventy percent of U.S. adults drink alcohol.
Suppose 5 U.S. adults are randomly selected. Let represent the number of adults in the sample who drink 
alcohol. Use the binomial probability formula, the binomial probability table, or your calculator to find the 
following probabilities.
\begin{itemize}
	\item a. That exactly 2 adults in the sample drink alcohol.
	= 0.1323
	\item b. That at least three adults in the sample drink alcohol.
	= P(3) + P(4) + P(5) = 0.3087 + 0.36015 + 0.16807 = 0.83692
	\item Alternatively, you can use binomcdf on the calculator:
	P(at least 3) = 1 – P(2 or fewer) = 1 – binomcdf(5, 0.70, 2) = 0.83692
	\item c. That everyone in the sample drinks alcohol.
	= 0.16807
\end{itemize}


%----------------------------------------------%

% PMS SPring 2006 Question 5

\section{Discrete Probability Distributions}
%==============================================================================%
\subsection{Confidence Intervals for Proportions}

\begin{itemize}
	\item In a survey conducted by a mail order company a random sample of 200 customers yielded 172 who indicated that they 
	were highly satisfied with the delivery time of their orders. 
	
	\item Calculate an approximate 95\% confidence interval for the proportion of the company's customers who are 
	highly satisfied with delivery times.
\end{itemize}


\[p= \frac{172}{200}= 86\%\]


\[ \frac{p(100-p)}{n} =\frac{86 \times 14}{200}\]









\section{Poisson Approximation}

n  = 25
p = 0.1, 0.2

Poisson approximation of Binomial ( letting $m=np$

\begin{itemize}
	\item $m_1 = 2.5$
	\item $m_2 = 5$
\end{itemize}

Find $P(X\geq 5)$ 

$ P(X\leq 4)  = 1 - P(X\geq 5) $

From Tables 
0.89118
0.44049

(Rest : Compare to Real Answers)

%========================================================%
% PMS Spring 2006 Question 3

A random sample of 10 is taken from a normal distribution of $\mu=20$ and $\sigma^2=1$. Let $s^2$ be the sample variance.

Find $P(S^2>1)$

\subsection*{Solution}

\[ \frac{(n-1)S^2}{\sigma^2} \sim \chi^2_{n-1}\]

\[ 9S^2 \sim \chi^2_{9}\]


\[ P(S^2>1) = P(\chi^2_{9}>9) = 1.05627 = 0.437\]


\end{document}
