\documentclass[12pt]{article}
\usepackage{amsmath}
%\usepackage[paperwidth=21cm, paperheight=29.8cm]{geometry}
\usepackage[angle=0,scale=1,color=black,hshift=-0.4cm,vshift=15cm]{background}
\usepackage{multirow}
\usepackage{enumerate}
\usepackage[gen]{eurosym}

%\SetBgScale{1}
%\SetBgAngle{0}
%\SetBgColor{black}
%\SetBgContents{\rule{1pt}{30cm}}
%\SetBgHshift{-8.4cm}
%
%\backgroundsetup{contents={
%\begin{tabular}{c|c}
%\hspace{2cm} & \\[0.7cm]
%& {\bf Statistics for Computing ------ Lecture 1 ------ Solutions} \\[0.3cm]
%%\hline
%\hspace{2cm} & \hspace{18.5cm} \\ [28cm]
%\end{tabular}}}

\backgroundsetup{contents={
{\bf \centering Statistics for Computing ------------------ Tutorial 7 --------------------------- Questions} }}


\setlength{\voffset}{-3cm}
\setlength{\hoffset}{-2cm}
\setlength{\parindent}{0cm}
\setlength{\textheight}{27cm}
\setlength{\textwidth}{17cm}

\pagestyle{empty}



\begin{document}

\subsection*{Question 1}
Assume that a character in a game is programmed to have an attack power according to $X \sim \text{Normal}(\mu=40,\sigma=3)$.\\[-0.2cm]

{\bf(a)} What is the probability that the attack is greater than 45? \quad {\bf(b)} What is the probability that the attack is between 32 and 42? \quad {\bf(c)} Let $X_1$ and $X_2$ be the first and second attacks. What is the probability that the \emph{sum} of these two attacks is greater than 85 units? \quad \\{\bf(d)} Calculate 99\% limits for the sum of two attacks.  \quad {\bf(e)} What is the probability that the \emph{difference} in attacks is more than 5 units? Note that attack 2 can be 5 units more than attack 1 or attack 1 can be 5 units more than attack 2, i.e., $\Pr(|D|>5)=\Pr(D<-5) + \Pr(D>5)$.



\subsection*{Question 2}

A character in a game deals a standard attack 75\% of the time and a critical attack the rest of the time (call these events $S$ and $S^c$). Given that it is a standard attack, the attack power is $X\,|\,S \sim \text{Normal}(\mu=40,\sigma=3)$. When the character deals a critical attack, a random fluctuation is added to this according to a $\text{Normal}(\mu=5,\sigma=1)$ distribution.\\[-0.2cm]

{\bf(a)} What is the distribution of $X\,|\,S^c$? \quad {\bf(b)} Calculate $\Pr(X<43\,|\,S)$ and $\Pr(X<43\,|\,S^c)$. \quad {\bf(c)} Calculate $\Pr(X<43)$. \,\, (hint: law of total probability) \quad {\bf(d)} If the character deals less than 43 damage points, what is the probability that the attack was a critical attack?



\subsection*{Question 3}
The income of a technician (in thousands) is $X_1 \sim \text{Normal}(\mu=30,\sigma=2)$. The income of an engineer is $X_2 \sim \text{Normal}(\mu=40,\sigma=3.5)$. \\[-0.2cm]

{\bf(a)} Calculate the probability that an engineer earns more than a technician. \quad {\bf(b)} Calculate 90\% limits for the difference in their income. \quad {\bf(c)} For a group of 25 technicians, calculate the probability that the average wage is less than 30500, i.e., $\Pr(\,\overline{\!X}_1 < 30.5)$. \quad {\bf(d)} In a group of 10 engineers, what is the probability that \emph{at least two} of them earn more than 45000? (hint: binomial with $p = \Pr(X_2 > 45)$) \quad {\bf(e)} For a sample of 30 technicians and 35 engineers, calculate the 80\% limits for the difference in their average wages.


\subsection*{Question 4}
Let $X \sim \text{Exponential}(\lambda=0.02)$. Calculate the following:\\[-0.2cm]

{\bf(a)} $\Pr(\,\overline{\!X} > 55)$ in a group of 100. \quad {\bf(b)} $\Pr(\,\overline{\!X} < 53)$ in a group of 40. \quad {\bf(c)} The value of $\bar x$ such that $\Pr(\,\overline{\!X} > \bar x) = 0.1$ when $n=65$. \quad {\bf(c)} The value of $n$ if $\Pr(\,\overline{\!X} < 49) = 0.1$.








\end{document} 