\documentclass[12pt]{article}
\usepackage{amsmath}
%\usepackage[paperwidth=21cm, paperheight=29.8cm]{geometry}
\usepackage[angle=0,scale=1,color=black,hshift=-0.4cm,vshift=15cm]{background}
\usepackage{multirow}
\usepackage{enumerate}

%\SetBgScale{1}
%\SetBgAngle{0}
%\SetBgColor{black}
%\SetBgContents{\rule{1pt}{30cm}}
%\SetBgHshift{-8.4cm}
%
%\backgroundsetup{contents={
%\begin{tabular}{c|c}
%\hspace{2cm} & \\[0.7cm]
%& {\bf Statistics for Computing ------ Lecture 1 ------ Solutions} \\[0.3cm]
%%\hline
%\hspace{2cm} & \hspace{18.5cm} \\ [28cm]
%\end{tabular}}}

\backgroundsetup{contents={
{\bf \centering Statistics for Computing ------------------ Tutorial 1 --------------------------- Questions} }}


\setlength{\voffset}{-3cm}
\setlength{\hoffset}{-2cm}
\setlength{\parindent}{0cm}
\setlength{\textheight}{27cm}
\setlength{\textwidth}{17cm}

\pagestyle{empty}



\begin{document}

\subsection*{Question 1}
A researcher wants to know if the Irish public are in support of a new government policy - she believes that less than 30\% of the public support it. She wishes to test this hypothesis by carrying out a survey. From a list of all individuals living in her local area, shes draws a random sample of 80 individuals and finds that 23 are in favour.\\[-0.2cm]

{\bf(a)} What is the population?\quad {\bf(b)} What is the sample?\quad {\bf(c)} What is the value of $n$?\quad \\{\bf(d)} Identify the the parameter and statistic - give the symbol and value for both.\quad {\bf(e)} What type of data was collected on each of the individuals in the sample? \quad {\bf(f)}  Identify any potential bias.\quad {\bf(g)} Can we confirm her hypothesis?

\subsection*{Question 2}
Employees of a software firm were each assigned the task of developing a tool. Once completed, they each submitted their code and executable file. A number of details were recorded:\\[-0.2cm]

{\bf(a)} The number of lines of code.\quad {\bf(b)} Time taken to develop the tool. \quad {\bf(c)} Experience level of the employee (high, average, low). \quad {\bf(d)} The size of the executable file (megabytes).\quad \\ {\bf(e)} Does the tool work? (yes/no) \quad {\bf(f)}  The employee's age.\quad {\bf(g)} Gender.\\[-0.2cm]

For each of the above, identify the data type.

\subsection*{Question 3}

\begin{center}
\begin{tabular}{|cccccccccc|}
\hline
&&&&&&&&&\\[-0.4cm]
2 & 4 & 2 & 1 & 5 & 3 & 0 & 4 & 1 & 8 \\
\hline
\end{tabular}
\end{center}

For the above sample of numbers, calculate the following:\\[0.2cm]
{\bf(a)} Mean. \quad {\bf(b)} Median. \quad {\bf(c)} Standard deviation. \quad {\bf(d)} Inter-quartile range.




\subsection*{Question 4}
A laptop manufacturer wishes to test a particular brand of CPU. A sample of 30 CPUs were selected and used to perform an intensive task for 1 hour. The temperature of each one was then recorded.
\begin{center}
\begin{tabular}{|cccccccccc|}
\hline
&&&&&&&&&\\[-0.4cm]
29.7 & 34.6 & 34.8 & 35.1 & 35.9 & 36.0 & 36.7 & 36.8 & 37.6 & 37.9\\
38.1 & 38.2 & 38.5 & 38.7 & 39.3 & 39.6 & 40.1 & 40.1 & 40.3 & 40.3\\
41.0 & 41.1 & 41.4 & 41.6 & 42.2 & 42.2 & 43.0 & 44.5 & 44.5 & 47.9 \\
\hline
\end{tabular}
\end{center}

{\bf(a)} Construct a frequency table with 5 classes (use 29 as the first breakpoint). \quad {\bf(b)} Draw the histogram (use relative frequency). \quad {\bf(c)} Calculate the median. \quad {\bf(d)} Calculate the quartiles. \quad {\bf(e)} Using the lower/upper fences, identify any outliers. \quad {\bf(f)} Draw the boxplot. \quad {\bf(g)} Is the data symmetric, left-skewed or right-skewed?


\end{document} 