\documentclass[12pt]{article}
\usepackage{amsmath}
%\usepackage[paperwidth=21cm, paperheight=29.8cm]{geometry}
\usepackage[angle=0,scale=1,color=black,hshift=-0.4cm,vshift=15cm]{background}
\usepackage{multirow}
\usepackage{enumerate}
\usepackage[gen]{eurosym}

%\SetBgScale{1}
%\SetBgAngle{0}
%\SetBgColor{black}
%\SetBgContents{\rule{1pt}{30cm}}
%\SetBgHshift{-8.4cm}
%
%\backgroundsetup{contents={
%\begin{tabular}{c|c}
%\hspace{2cm} & \\[0.7cm]
%& {\bf Statistics for Computing ------ Lecture 1 ------ Solutions} \\[0.3cm]
%%\hline
%\hspace{2cm} & \hspace{18.5cm} \\ [28cm]
%\end{tabular}}}

\backgroundsetup{contents={
{\bf \centering Statistics for Computing ------------------ Tutorial 10 --------------------------- Questions} }}


\setlength{\voffset}{-3cm}
\setlength{\hoffset}{-2cm}
\setlength{\parindent}{0cm}
\setlength{\textheight}{27cm}
\setlength{\textwidth}{17cm}

\pagestyle{empty}



\begin{document}


\subsection*{Question 1}
A machine is set to produce laptop sleeves with the following dimensions:\\ length $\times$ width $\times$ depth = $40 \text{ cm} \times  30 \text{ cm} \times 2 \text{ cm}$. A sample of 40 sleeves was selected and each was measured. The results were as follows:\\[-0.2cm]
\begin{center}
\begin{tabular}{|c|c|c|c|}
\hline
&&&\\[-0.4cm]
& length & width & depth \\
\hline
&&&\\[-0.4cm]
$\bar x$ & 40.11 & 30.09  & 1.91 \\
\hline
&&&\\[-0.4cm]
$s$ & 0.51 & 0.17 & 0.15 \\
\hline
\end{tabular}
\end{center}

Test the following hypotheses (use the 5\% level of significance in each case):\\[0.2cm]
{\bf(a)} The mean length is equal to 40cm. \quad {\bf(b)} The mean width is equal to 30cm. \quad \\{\bf(c)} The mean depth is equal to 2cm. \quad {\bf(d)} Both the width and the depth of the sleeve need to be addressed here - which do you think is more urgent?  \quad {\bf(d)} Calculate the p-values for the tests carried out in parts (a), (b) and (c).



\subsection*{Question 2}
A matchbox is supposed to contain 100 matches. We wish to test this hypothesis.  \\[-0.2cm]

{\bf(a)} State the null and alternative hypotheses. \quad {\bf(b)} From a sample of 32 matchboxes, it is found that the average is 99.4 and the standard deviation is 2.1. Calculate the test statistic. \quad {\bf(c)} Provide your conclusion based on the p-value.



\subsection*{Question 3}
An aircraft part is designed to last more than 500 hours. However, in the interest of safety, it will first be assumed that the part lasts \emph{less than or equal to} 500 hours (i.e., this is the null hypothesis) unless there is firm evidence suggesting otherwise.\\[-0.2cm]

{\bf(a)} State the null and alternative hypotheses. \quad {\bf(b)} What is the critical value if $\alpha=0.001$ and only 4 units will be run until wearout (due to the expense of wasting aircraft parts). \quad \\ {\bf(c)} In this sample of size 4, it is found that the average is 566 hours and the variance is 83 hours$^2$. Calculate the test statistic. \quad {\bf(d)} What is the conclusion?



\subsection*{Question 4}
A friend claims that he can pass a particular game in 4 hours or less (on average). We wish to test this hypothesis at the 10\% level of significance. Your friend plays the game on 6 different occasions: his average completion time is 4.6 hours and the standard deviation is 0.5 hours. \\[-0.2cm]

{\bf(a)} State the null and alternative hypotheses. \quad {\bf(b)} What is the critical value? \quad {\bf(c)} Calculate the test statistic and provide your conclusion. \quad {\bf(d)} Between what two values does the p-value lie? (note: the p-value cannot be calculated exactly using the t-tables)



\subsection*{Question 5}
A die is rolled 80 times and we count 18 sixes. We wish to test the hypothesis that the die is fair (note: if this is the case, the proportion of sixes is $p = \frac{1}{6}$). \\[-0.2cm]

{\bf(a)} State the null and alternative hypotheses. \quad {\bf(b)} If we wish to test at the 5\% level of significance, what is the critical value? \quad {\bf(c)} Calculate the test statistic and provide your conclusion.



\subsection*{Question 6}
Assume that a particular brand dominates the market. More specifically, it is well-known that at least 60\% of people use this brand (i.e., $p \ge 0.6$). However, in response to recent media claims that this brand is weakening, the company wish to test the hypothesis that $p \ge 0.6$. \\[-0.2cm]

{\bf(a)} State the null and alternative hypotheses. \quad {\bf(b)} From a sample of 1000 people, it is found that 629 use this brand; calculate the test statistic and, hence, the p-value. \quad {\bf(c)} Based on the evidence, state your conclusion.



\subsection*{Question 7}
Last year 30\% of applicants to a graduate programme failed the aptitude test. This year 100 graduates applied - 25\% of these failed the test.\\[-0.2cm]

{\bf(a)} We wish to test the hypothesis that the quality of applicants has not changed since last year - what are the null and alternative hypotheses? \quad {\bf(b)} If we are testing at the 1\% level, what is the rejection region? \quad {\bf(c)} Based on the data, calculate the test statistic and provide your conclusion.





\subsection*{Question 8}
A soft drinks company is working on a new recipe for its best-selling drink. The company intends to carry out a study where participants will taste both flavours (current and new) and then answer the question:
    \begin{quotation}
    ``Do you prefer the new flavour?''
    \end{quotation}
    It is assumed that the \emph{current} recipe is superior, i.e., that \emph{less than or equal to} 50\% of people prefer the new drink ($p \le 0.5$).\\[0.4cm]
    We wish to test the hypothesis that $p \le 0.5$.\\[-0.2cm]

{\bf(a)} State the null and alternative hypotheses. \quad {\bf(b)} From a sample of 65 people, we find that 43 people prefer the new recipe. Calculate the test statistic and, hence, the p-value. \quad \\ {\bf(c)} Based on the evidence, state your conclusion.










\end{document} 